Contextuality and Nonlocality in Discrete Quantum Theory

### 2016-02-10 DQC3.tex
"6.1 Gleason's Theorem in DQT" Updates new results...

If the problem is that the range of $\hat{\pi}$ is only a commutative
semigroup, we may want to consider $\hat{\pi}:\ffzd{p^{2}}\rightarrow\mathscr{L}$
, and $\left(\mathscr{L},0,\uplus\right)$ is an abelian group. By
the fundamental theorem of abelian group~\cite{DummitFoote2004},
$\mathscr{L}$ must be isomorphic to $\mathbb{Z}_{d_{1}}\oplus\ldots\oplus\mathbb{Z}_{d_{k}}\oplus\mathbb{Z}^{b}$,
where $\mathbb{Z}_{d_{1}},\ldots,\mathbb{Z}_{d_{k}}$ are finite cyclic
subgroups, $\mathbb{Z}^{b}$ is the free abelian group with rank $b$,
and $\oplus$ is the direct sum of the abelian groups. Moreover, we
want $\left(\mathscr{L},0,\uplus\right)$ to have a sense of probability. 
\begin{problem}
\label{thm:GleasonDQT-group}Is there a group $\left(\mathscr{L},0,\oplus\right)$
with a special element $1\in\mathscr{L}$ and $1\ne0$ satisfying
the following statement? 
\begin{itemize}
\item There is a way to define a mixed state $\rho$ and a corresponding
Born rule $\tilde{\pi}_{\rho}:\ffzd{p^{2}}\rightarrow\mathscr{L}$
in a sense of equation (\ref{prob.mixed}).
\item For $d\ge3$, for any $\hat{\pi}:\ffzd{p^{2}}\rightarrow\mathscr{L}$
satisfying 
\begin{eqnarray}
 &  & \textrm{For any vector }\ket{i}\in\ffzd{p^{2}}\textrm{, and }\gamma\in\ff{p^{2}}^{*}\textrm{ , }\hat{\pi}(\ket{i})=\hat{\pi}(\gamma\ket{i})\textrm{ . }\label{eq:parallel-DQT-1}\\
 &  & \textrm{For any orthonormal basis }\{\ket{i}\}\textrm{ , }{\displaystyle \bigoplus_{i=0}^{d-1}\hat{\pi}\left(\ket{i}\right)=}1\textrm{ .}\label{eq:total-probability-DQT-1}
\end{eqnarray}
There is a mixed state $\rho$ such that $\hat{\pi}=\tilde{\pi}_{\rho}$.
\end{itemize}
\end{problem}
When $d=3$ and $p=7$, we have verified that we no longer have the
freedom to assign impossible more than any reasonable probability
assignment. Therefore, we cannot find a similar counterexample for
problem \ref{thm:GleasonDQT-group}. However, even if there is a group
$\left(\mathscr{L},0,\oplus\right)$ satisfying problem \ref{thm:GleasonDQT-group},
the corresponding Born rule $\tilde{\pi}_{\rho}:\ffzd{p^{2}}\rightarrow\mathscr{L}$
is not basis independent, i.e., we will not have 
\begin{equation}
\tilde{\pi}_{U\rho}\left(U\ket{i}\right)=\tilde{\pi}_{\rho}\left(\ket{i}\right)\label{eq:rotation-1}
\end{equation}
for all unitary matrix $U$, mixed state $\rho$, and pure state $\ket{i}\in\ffzd{p^{2}}$,
and the proof is similar to the one of the no-go theorem in section
\ref{sub:A-No-Go-Theorem}.

In summary, there are some difficulties to establish corresponding
Gleason's theorem in DQT no matter how we set the conditions. However,
Gleason's theorem gives an idea how to fine ``the'' reasonable Born
rule in a discrete quantum theory, and we hope we may find one in
the future.

>>>>

If the problem is that the range of $\hat{\pi}$ is only a commutative
semigroup, we may want to consider $\hat{\pi}:\ffzd{p^{2}}\rightarrow\mathscr{L}$
, and $\left(\mathscr{L},0,\oplus\right)$ is an abelian group. By
the fundamental theorem of finitely generated abelian group~\cite{DummitFoote2004},
$\mathscr{L}$ must be isomorphic to $\mathbb{Z}_{d_{1}}\times\ldots\times\mathbb{Z}_{d_{k}}\times\mathbb{Z}^{b}$,
where $\mathbb{Z}_{d_{1}},\ldots,\mathbb{Z}_{d_{k}}$ are finite cyclic
subgroups, $\mathbb{Z}^{b}$ is the free abelian group with rank $b$,
and $\times$ is the direct product of the abelian groups. However,
if $\mathscr{L}$ has any finite cyclic subgroups, we may have $\hat{\pi}\left(\ket{0}\right)\ne0$,
but $d_{1}\hat{\pi}\left(\ket{0}\right)=0$. This contradicts the
idea that $\hat{\pi}$ can be understood as a probability. Therefore,
we must have $\mathscr{L}\cong\mathbb{Z}^{b}$.
\begin{problem}
\label{thm:GleasonDQT-group}Is there a group $\left(\mathscr{L},0,\oplus\right)\cong\mathbb{Z}^{b}$
with a special element $1\in\mathscr{L}$ and $1\ne0$ satisfying
the following statement? 
\begin{itemize}
\item There is a way to define a mixed state $\rho$ and a corresponding
Born rule $\tilde{\pi}_{\rho}:\ffzd{p^{2}}\rightarrow\mathscr{L}$
in a sense of equation (\ref{prob.mixed}).
\item $\hat{\pi}:\ffzd{p^{2}}\rightarrow\mathscr{L}$ satisfies 
\begin{eqnarray}
 &  & \hat{\pi}(\ket{i})=\hat{\pi}(\gamma\ket{i})\textrm{ for any vector }\ket{i}\in\ffzd{p^{2}}\textrm{, and }\gamma\in\ff{p^{2}}^{*}\textrm{ , }\label{eq:parallel-DQT-1}\\
 &  & {\displaystyle \bigoplus_{i=0}^{d-1}\hat{\pi}\left(\ket{i}\right)=}1\textrm{ for any orthonormal basis }\{\ket{i}\}\textrm{ . }\label{eq:total-probability-DQT-1}
\end{eqnarray}
For $d\ge3$, there is a mixed state $\rho$ such that $\hat{\pi}=\tilde{\pi}_{\rho}$;
for $d=2$, there is a $\hat{\pi}$ corresponding to no mixed state
$\rho$.
\end{itemize}
\end{problem}
For $d=2$, we can easily construct a $\hat{\pi}$ satisfying equation
(\ref{eq:zero-count}) and corresponding to no mixed state $\rho$.
When $d=3$ and $p=7$, we have verified the only $\hat{\pi}$ is
$\hat{\pi}\left(\ket{i}\right)=\hat{\pi}\left(\ket{j}\right)$ and
$3\hat{\pi}\left(\ket{i}\right)=1$ for all $\ket{i},\ket{j}\in\ffzd{p^{2}}$.
Therefore, we don't have enough $\hat{\pi}$ to corresponding to even
pure states. We believe the same situation happens for other $d\ge3$
and $p\ge7$.

We don't know whether $\mathscr{L}$ can be a non-Boolean commutative
semigroup, but not yet a group. In summary, there are some difficulties
to establish corresponding Gleason's theorem in DQT no matter how
we set the conditions.
